Are Transformed Low Discrepancy Points Also Low Discrepancy?

### 2019-06-02 InvPts_verB.tex

\usepackage{psfrag,bbm,epsf}
\usepackage{xspace,subfigure}
\usepackage{amssymb, xcolor,enumerate,float,mathtools}
\usepackage{algpseudocode}
\usepackage{algorithm, algorithmicx}
\usepackage[numbers]{natbib}
\usepackage{enumerate,multirow,booktabs}
\usepackage{url}
\usepackage{natbib}
\usepackage{showlabels}

@@

\emph{Example}: Consider estimating the integral
\begin{align}\label{eq:badexample}
\int_{\mathbb{R}^d} \frac{x_1+\cdots+x_d}{1+10^(-8)(x_1^2+\cdots+x_d^2)}\dif \Phi(\vx) & = \Ex(e^{|x_1+\cdots+x_d|}),\,\,\,\vx\sim N({\bf 0}, {\mathsf I}_d)
\end{align}
using Monte Carlo method, with $d=10$ and $N=2^{10}$ points. We start with two uniform random sequences. One is a Sobol sequence that has small uniform discrepancy, denoted as $\Udes_1$. Then we move the closest point to origin $0^d$ in $\Udes_1$ to $\left(10^{-15}\right)^d$, and denote the new sequence as $\Udes_2$. We compare the integral estimate with samples $\vPsi(\Udes_1)$ and $\vPsi(\Udes_2)$. Table \ref{tab:badexample} shows that the uniform discrepancies of $\Udes_1$ and $\Udes_2$ don't differ much, but the normal discrepancy of $\vPsi(\Udes_2)$ is far larger than that of $\vPsi(\Udes_1)$. What's more, the integral estimate using $\vPsi(\Udes_2)$ explodes while that of $\vPsi(\Udes_1)$ is reasonable.

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
  \centering
  \caption{Comparison of Integral Estimate}
    \begin{tabular}{ccccc}
    \hline
   $\Udes$       & $D(\Udes,F_\unif, K)$     & $D(\vPsi(\Udes),F_\normal, K)$     & Estimate \\
    \hline
    $\Udes_1$     &   0.0286    &  18.57     & 251.33 \\
    $\Udes_2$     &  0.0294      &  58.82     & $3.01\times 10^{31}$ \\
    \hline
    \end{tabular}%
  \label{tab:badexample}%
\end{table}%

@@

To further evaluate the performance of the normal random sequence generated by the four generators, we consider the integral approximation problem in \eqref{eq:badexample} with $d=10$. We use the Monte Carlo estimate with the sample generated by SOBOL of size $n = 2^{17}$ as an accurate estimate of the integral, and summarize the integral estimation error according to the four generators with simulation size $N=512$ in Figure \ref{fig:comparison_integral}. Not surprisingly, CE outperforms the other three generators and has no explosive relative error. It's interesting to see that SOBOL and E-SOBOL may perform even worse than RAND in several cases, and this aligns with our discussion in Section 4: the inverse transformed low uniform discrepancy sequence may not retain low normal discrepancy. When the sample size $N$ is small/moderate relative to dimension $d$, using CE would limit the potential danger of very inaccurate estimate caused by a sample with large normal discrepancy.

\begin{figure}[ht]\label{fig:comparison_integral}
\centering
{\subfigure[$d=2,N=32$] {\includegraphics[width=5.5cm]{code/errorboxplot_d2n32.eps}}}
\quad
{\subfigure [$d=3, N=64$]
{\includegraphics[width=5.5cm]{code/errorboxplot_d3n64.eps}}}
\quad
{\subfigure[$d=4, N=64$] {\includegraphics[width=5.5cm]{code/errorboxplot_d4n64.eps}}}
\quad
{\subfigure[$d=6, N=128$] {\includegraphics[width=5.5cm]{code/errorboxplot_d6n128.eps}}}
\quad
{\subfigure[$d=8, N=256$] {\includegraphics[width=5.5cm]{code/errorboxplot_d8n256.eps}}}
\quad
{\subfigure[$d=10, N=512$] {\includegraphics[width=5.5cm]{code/errorboxplot_d10n512.eps}}}
\caption{Integral Approximate Performance Comparison of Generators}
\end{figure}

>>>>

\usepackage{psfrag,bbm,epsf}
\usepackage{xspace,subfigure}
\usepackage{comment}
\usepackage{amssymb, xcolor,enumerate,float,mathtools}
\usepackage{algpseudocode}
\usepackage{algorithm, algorithmicx}
\usepackage[numbers]{natbib}
\usepackage{enumerate,multirow,booktabs}
\usepackage{url}
\usepackage{natbib}
\usepackage{showlabels}

@@

\emph{Example}: Consider estimating the integral
\begin{align}\label{eq:badexample}
\int_{\mathbb{R}^d} \frac{x_1+\cdots+x_d}{1+10^{-8}(x_1^2+\cdots+x_d^2)}\dif \Phi(\vx) & = \Ex\left[\frac{\sum\limits_{i=1}^d x_i}{1+10^{-8}\|\vx\|_2^2}\right],\,\,\,\vx\sim N({\bf 0}, {\mathsf I}_d)
\end{align}
using Monte Carlo method, with $d=10$ and $N=2^{10}$ points. Analytically, the integral \eqref{eq:badexample} equals 0 since $\frac{x_1+\cdots+x_d}{1+10^{-8}(x_1^2+\cdots+x_d^2)}$ is an odd function in each dimension and the probability density function of standard normal is an even function. We start with two uniform random sequences. One is a Sobol sequence that has small uniform discrepancy, denoted as $\Udes_1$. Then we move the closest point to origin $0^d$ in $\Udes_1$ to $\left(10^{-15}\right)^d$, and denote the new sequence as $\Udes_2$. We compare the integral estimate with samples $\vPsi(\Udes_1)$ and $\vPsi(\Udes_2)$. Table \ref{tab:badexample} shows that the uniform discrepancies of $\Udes_1$ and $\Udes_2$ don't differ much, but the normal discrepancy of $\vPsi(\Udes_2)$ is far larger than that of $\vPsi(\Udes_1)$. In return, the estimate error corresponding to $\Udes_2$ is about 60 times larger than that corresponding to $\Udes_1$.

% Table generated by Excel2LaTeX from sheet 'Sheet1'
\begin{table}[htbp]
  \centering
  \caption{Comparison of Integral Estimate}
    \begin{tabular}{ccccc}
    \hline
   $\Udes$       & $D(\Udes,F_\unif, K)$     & $D(\vPsi(\Udes),F_\normal, K)$     & Estimate \\
    \hline
    $\Udes_1$     &   0.0285    &  18.57     & 0.0011 \\
    $\Udes_2$     &  0.0292      &  58.82     & -0.0660 \\
    \hline
    \end{tabular}%
  \label{tab:badexample}%
\end{table}%

@@

\begin{comment}
To further evaluate the performance of the normal random sequence generated by the four generators, we consider the integral approximation problem in \eqref{eq:badexample} with $d=10$. We use the Monte Carlo estimate with the sample generated by SOBOL of size $n = 2^{17}$ as an accurate estimate of the integral, and summarize the integral estimation error according to the four generators with simulation size $N=512$ in Figure \ref{fig:comparison_integral}. Not surprisingly, CE outperforms the other three generators and has no explosive relative error. It's interesting to see that SOBOL and E-SOBOL may perform even worse than RAND in several cases, and this aligns with our discussion in Section 4: the inverse transformed low uniform discrepancy sequence may not retain low normal discrepancy. When the sample size $N$ is small/moderate relative to dimension $d$, using CE would limit the potential danger of very inaccurate estimate caused by a sample with large normal discrepancy.

\begin{figure}[ht]\label{fig:comparison_integral}
\centering
{\subfigure[$d=2,N=32$] {\includegraphics[width=5.5cm]{code/errorboxplot_d2n32.eps}}}
\quad
{\subfigure [$d=3, N=64$]
{\includegraphics[width=5.5cm]{code/errorboxplot_d3n64.eps}}}
\quad
{\subfigure[$d=4, N=64$] {\includegraphics[width=5.5cm]{code/errorboxplot_d4n64.eps}}}
\quad
{\subfigure[$d=6, N=128$] {\includegraphics[width=5.5cm]{code/errorboxplot_d6n128.eps}}}
\quad
{\subfigure[$d=8, N=256$] {\includegraphics[width=5.5cm]{code/errorboxplot_d8n256.eps}}}
\quad
{\subfigure[$d=10, N=512$] {\includegraphics[width=5.5cm]{code/errorboxplot_d10n512.eps}}}
\caption{Integral Approximate Performance Comparison of Generators}
\end{figure}
\end{comment}
